---
title: "Time Series Analysis of Fish Passage in Willamette Falls"
author: "Kiera Matiska"
date: "3/5/2022"
output: 
  html_document: 
    code_folding: hide
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# attach packages
library(tidyverse)
library(here)
library(broom)
library(janitor)
library(lubridate)
library(tsibble)
library(feasts)
library(kableExtra)
library(patchwork)
```

## Read in Data

```{r}
fish <- read_csv(here("data", "willamette_fish_passage.csv")) %>% 
  clean_names()

fish_NA_0 <- replace(fish, is.na(fish), 0)

fish_ts <- fish_NA_0 %>% 
  mutate(date = mdy(date)) %>% 
  as_tsibble(key = NULL, index = date) %>% 
  select(project, date, coho, jack_coho, steelhead)
```

## Time Series Visualizations {.tabset .tabset-fade}

### Original Time Series

```{r}
coho_plot <- ggplot(data = fish_ts,
       aes(x = date, y = coho)) +
  geom_line(color = "dodgerblue3") + 
  labs(title = "Daily Coho Salmon \nRecords in Willamette \nFalls from 2000-2010",
       x = "Date",
       y = "Individual Count") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "aliceblue"))

jack_plot <- ggplot(data = fish_ts,
       aes(x = date, y = jack_coho)) +
  geom_line(color = "dodgerblue3") +
  labs(title = "Daily Jack Coho Salmon \nRecords in Willamette \nFalls from 2000-2010",
       x = "Date",
       y = "Individual Count") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "aliceblue"))

steelhead_plot <- ggplot(data = fish_ts,
       aes(x = date, y = steelhead)) +
  geom_line(color = "dodgerblue3") +
  labs(title = "Daily Steelhead Trout \nRecords in Willamette \nFalls from 2000-2010",
       x = "Date",
       y = "Individual Count") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "aliceblue"))

wrap_plots(coho_plot, jack_plot, steelhead_plot)
```

**Figure 1.** *In order from left to right:* (A) Daily Coho Salmon Records in Willamette Falls from 2000-2010; (B) Daily Jack Coho Salmon Records in Willamette Falls from 2000-2010; and (C) Daily Steelhead Trout Records in Willamette Falls from 2000-2010. These graphs illustrate the pattern of ladder use of each species during this 10-year time period.

#### Summary

- Both Coho And Jack Coho Salmon use the Fish Ladder later in the year; Steelhead Trout use it more consistently throughout the year.
- Between these years, it appears that the Coho Salmon population has increased substantially based on the amount of individuals going through the fish ladder.
- Between the years of 2006 an 2010 for Steelhead Trout, there was a substantial decrease in ladder use than in previous years by the species.

### Seasonplots

```{r}
coho_season <- fish_ts %>% 
  gg_season(y = coho) +
  labs(title = "Seasonal Records of Coho \nSalmon in Willamette \nFalls from 2000-2010",
       x = "Month",
       y = "Individual Count") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "aliceblue"),
        legend.position = "bottom") +
  guides(color = guide_legend(nrow = 4))

jack_season <- fish_ts %>% 
  gg_season(y = jack_coho) +
  labs(title = "Seasonal Records of Jack \nCoho Salmon in Willamette \nFalls from 2000-2010",
       x = "Month",
       y = "Individual Count") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "aliceblue"),
        legend.position = "bottom") +
  guides(color = guide_legend(nrow = 4))

steelhead_season <- fish_ts %>% 
  gg_season(y = steelhead) +
  labs(title = "Seasonal Records of Steelhead \nTrout in Willamette \nFalls from 2000-2010",
       x = "Month",
       y = "Individual Count") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "aliceblue"),
        legend.position = "bottom") +
  guides(color = guide_legend(nrow = 4))

wrap_plots(coho_season, jack_season, steelhead_season)
```

**Figure 2.** *From left to right:* (A) Seasonal Records of Coho Salmon in Willamette Falls from 2000-2010; (B) Seasonal Records of Jack Coho Salmon in Willamette Falls from 2000-2010; and (C) Seasonal Records of Steelhead Trout in Willamette Falls from 2000-2010.

#### Summary

- Coho and Jack Coho Salmon are found using the ladder during autumn, with most recorded occurrences happen in October.
- Steelhead Trout are found to use the ladder between the months of January-July and December each year. Most use of the ladder by this species comes in late spring and summer, between the months of April and July.
- Steelhead Trout has the longest period of ladder use, followed by Coho Salmon. Jack Coho Salmon are found to be using the ladder the least.

### Annual Counts {.tabset .tabset-fade}

```{r}
fish_annual <- fish_ts %>% 
  index_by(yearly = ~year(.)) %>% 
  summarize(annual_coho = sum(coho),
            annual_jack_coho = sum(jack_coho),
            annual_steelhead = sum(steelhead))
```

#### Coho Salmon

```{r}
fish_annual_coho <- fish_annual %>% 
  select(yearly, annual_coho)

fish_annual_coho %>% 
  kbl()
```

#### Jack Coho Salmon

```{r}
fish_annual_jack <- fish_annual %>% 
  select(yearly, annual_jack_coho)

fish_annual_jack %>% 
  kbl()
```


#### Steelhead Trout

```{r}
fish_annual_steelhead <- fish_annual %>% 
  select(yearly, annual_steelhead)

fish_annual_steelhead %>% 
  kbl()
```

- A finalized figure of annual totals for fish passage, for each of the 3 species (coho, jack coho, steelhead). You decide if this is in a single panel, or multiple panels. Add a figure caption. 
- Add 2 - 3 bullet points summarizing major trends you see in the annual totals by species from 2000 - 2010.

